Constrain the Merger History of Primordial-Black-Hole Binaries from GWTC-3

### 2022-10-28 latex/paper.tex
update

%\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Primordial black holes (PBHs), as dark matter candidates, have attracted 
more and more attentions as they could be possible progenitors of 
the heavy binary black holes (BBHs) observed by \lvc.
Accurately estimating the merger rate of PBH binaries will be crucial to
reconstruct the mass distribution of PBHs.
%However there exists some uncertainties in estimating the merger rate.
It was pointed out the merger history of PBHs may shift the merger rate distribution depending on the mass function of PBHs.
In this paper, we use BBH events from \lvc\ GWTC-3 data release to
constrain the merger rate distribution of PBHs by taking into account the effect of merger history.
It is found that the second merger process makes subdominant contribution to the 
total merger rate, and hence the merger history effect can be safely neglected.	
We also confirm that the main components of dark matter should not be made of stellar 
mass PBHs.

%\keywords{primordial black holes, merger rate, merger history}
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
The direct detection of gravitational wave (GW) from a binary black hole (BBH)
coalescence \cite{Abbott:2016blz} has opened a new window into the universe.
Over the past few years, ten BBH mergers have been reported by \lvc\ during 
the O1 and O2 observing runs \cite{Abbott:2016blz,Abbott:2016nmj,%
TheLIGOScientific:2016pea,Abbott:2017vtc,%
Abbott:2017gyy,Abbott:2017oio,LIGOScientific:2018mvr}.
The progenitors of these BBHs are still unknown and under intensively investigation (see \eg \cite{Bird:2016dcv,Sasaki:2016jop,Chen:2018czv,%
Fishbach:2017dwv,Clesse:2016vqa,Antonini:2016gqe,%
Inayoshi:2017mrs,Ali-Haimoud:2017rtz,Perna:2019axr,Kavanagh:2018ggo,%
Rodriguez:2015oxa,Rodriguez:2016kxx,Park:2017zgj,%
Belczynski:2014iua,Belczynski:2016obo,Woosley:2016nnw,%
Rodriguez:2018rmd,Choksi:2018jnq,2010AIPC.1314..291D,deMink:2016vkw}).
These \lvc\ BBHs present a much heavier mass distribution (in particular the 
source-frame primary mass of GW170729 event can be as heavy as
$50.2^{+16.2}_{-10.2} \Msun$ \cite{LIGOScientific:2018mvr}) than that 
inferred from X-ray observations
\cite{Wiktorowicz:2013dua,Casares:2013tpa,Corral-Santana:2013uua,%
Corral-Santana:2015fud},
which would challenge the formation and evolution mechanisms of astrophysical
black holes. 
One possible explanation for \lvc\ BBHs is the primordial black holes (PBHs)
\cite{Bird:2016dcv,Sasaki:2016jop,Chen:2018czv}
formed through the gravitational collapse of the primordial density
fluctuations \cite{Hawking:1971ei,Carr:1974nx}, 
which may accompany the induced GWs
\cite{Yuan:2019udt,Yuan:2019wwo,Chen:2019xse,Yuan:2019fwv,Cai:2019bmk,Cai:2018dig,Cai:2019elf,DeLuca:2019ufz,Bartolo:2018rku,Bartolo:2018evs}.
On the other hand, PBHs can also be a candidate of cold dark matter (CDM),
and the abundance of PBHs in CDM has been constrained by various experiments
\cite{Carr:2009jm,Barnacka:2012bm,Graham:2015apa,Niikura:2017zjd,%
Griest:2013esa,Niikura:2019kqi,Tisserand:2006zx,Brandt:2016aco,Gaggero:2016dpq,%
Ali-Haimoud:2016mbv,Blum:2016cjs,Horowitz:2016lib,Chen:2016pud,%
Wang:2016ana,Abbott:2018oah,Magee:2018opb,Wang:2019kaf,Chen:2018rzo,Chen:2019irf,%
Yuan:2019udt,Chen:2019xse,Montero-Camacho:2019jte,Laha:2019ssq,Dasgupta:2019cae,Gow:2019pok}.

In order to account for the \lvc\ BBHs, the merger rate of PBH binaries has
been estimated to be $17 \sim 288~\gpcyr$ \cite{Chen:2018rzo}.
One should notice that theoretically there exist some uncertainties in
estimating the merger rate distribution of PBH binaries, and the accuracy of 
the estimation has been continuously improved.
The merger rate of PBH binaries with monochromatic mass function has already
been given in \cite{Nakamura:1997sm,Sasaki:2016jop} for the case where two 
neighboring PBHs having sufficiently small separation can form a binary in
the early Universe due to the torque from the third nearest PBH.
These binaries would then evolve and coalesce within the age of the Universe 
and finally explain the merger events observed by \lvc\
\cite{Sasaki:2016jop}.
Later, the merger rate estimation is improved in \cite{Ali-Haimoud:2017rtz} 
by taking into account the torques exerted by all CDM (including all the PBHs
and linear density perturbations), but it is also assumed that all PBHs have
the same mass.
It is also pointed out in \cite{Ali-Haimoud:2017rtz} that the effects such as
encountering with other PBHs, tidal field from the smooth halo, and
the baryon accretion are subdominant and can be neglected when estimating
the merger rate.

Various attempts have been made to estimate the merger rate distribution
of PBH binaries when PBHs have an extended mass function 
\cite{Chen:2018czv,Chen:2018rzo,Raidal:2018bbj,Raidal:2017mfl,Liu:2018ess,Liu:2020cds,Vaskonen:2019jpv}.
In particular, a formalism to estimate the effect of merger history of PBHs
on merger rate distribution has been developed in \cite{Liu:2019rnx},
and it is argued that the multiple-merger effect may not be ignored if
PBHs have a power-law or a log-normal mass function by choosing some specific
parameter values of the mass function.
An accurate estimation of the merger rate will be crucial to infer the event
rate of \lvc\ BBHs and constrain the abundance of PBHs in CDM either through
null detection of sub-solar mass PBH binaries or the null detection
of stochastic GW background (SGWB) from PBH binaries.

In this paper, we will use the LVK recent released GWTC-3 data to constrain the effect of merger history on the merger rate of PBH binaries assuming all of LVK BBHs are of primordial origin.
We organize the rest paper as follows.
In \Sec{merger}, we give a brief review on calculation of merger rate of PBH binaries by accounting for the merger history effect.
In \Sec{method}, we describe the hierarchical Bayesian framework used to infer the PBH population parameters from GW data.
In \Sec{result}, we consider four commonly used PBH mass functions and present the results.
Finally, we give conclusions in \Sec{conclusion}.

@@

In this section, we will outline the calculation of merger rate density when taking into account the PBH merger history effect. We refer to \cite{Liu:2019rnx} for more details.

The BBHs observed by LVK suggest that BHs should have a broad mass distribution, so we consider an extended mass function for PBHs. Here, we demand the probability distribution function of PBH mass, $P(m)$, be normalized such that
\e\label{norm}
    \int_{0}^{\infty} P(m)\, \rd m = 1.
\q 
Assuming the fraction of PBHs in CDM is $\fpbh$, we can estimate the abundance of PBHs in the mass interval $(m, m+\rd m)$ as \citep{Chen:2018rzo}
\e 
    0.85 \fpbh\, P(m)\, \rd m,
\q 
where the coefficient $0.85$ is roughly the fraction of CDM in non-relativistic matter including both CDM and baryons. Following \cite{Liu:2019rnx}, we may define an average PBH mass, $\mpbh$, as
\e\label{mpbh}
    \frac{1}{\mpbh} = \int \frac{P(m)}{m} \rd m.
\q 
Then, we can obtain the average number density of PBHs with mass $m$ in the 
total number density of PBHs, $F(m)$, by \citep{Liu:2019rnx}
\e\label{Fm} 
    F(m) = P(m) \frac{\mpbh}{m}.
\q

We can now proceed to estimate the merger rate densities of PBH binaries by considering the merger history effect.
We assume that PBHs are randomly distributed in the early Universe when they decouple from the cosmic background evolution \citep{Nakamura:1997sm,Sasaki:2016jop,Ali-Haimoud:2017rtz}. Two nearest PBHs would attract each other because of the gravitational interactions. These two PBHs would obtain the angular momentum from the torque of other PBHs and form a PBH binary. The binary would emit gravitational radiations and eventually merge.

We do not intend to give a detailed derivation, but quote the results from \cite{Liu:2019rnx} here.
The merger rate density from first-merger process, $\cR_1(t, m_i, m_j)$, is given by \citep{Liu:2019rnx}
\e 
\cR_1(t, m_i, m_j) = \int \hat{\cR}_1 \rd m_l,
\q 
where
\e 
\begin{split}
    &\hat{\cR}_1(t, m_i, m_j, m_l)
    \equiv 1.32 \times 10^6 \times \(\frac{t}{t_0}\)^{-\frac{34}{37}}\(\frac{\fpbh}{\mpbh}\)^\frac{53}{37} \\ 
    &\times m_l^{-\frac{21}{37}} (m_i m_j)^\frac{3}{37} (m_i + m_j)^\frac{36}{37} F(m_i) F(m_j) F(m_l).     
\end{split}
\q
Similarly, the merger rate density from second-merger process, $\cR_2(t, m_i, m_j)$, is given by \citep{Liu:2019rnx}
\e 
\begin{split}
    \cR_2(t, m_i, m_j) &= \hf \int \hat{\cR}_2(t, m_i-m_e, m_e, m_j, m_l)\ \rd m_l \rd m_e\\
    &+ \hf \int \hat{\cR}_2(t, m_j-m_e, m_e, m_i, m_l)\ \rd m_l \rd m_e,
\end{split}
\q 
where
\e 
\begin{split}
    &\hat{\cR}_2(t, m_i, m_j,m_k, m_l)
    =1.59 \times 10^4 \times \(\frac{t}{t_0}\)^{-\frac{31}{37}} \(\frac{\fpbh}{\mpbh}\)^\frac{69}{37}\\ 
    &\qquad\times m_k^{\frac{6}{37}} m_l^{-\frac{42}{37}} (m_i+ m_j)^\frac{6}{37} (m_i + m_j + m_k)^\frac{72}{37}\\
    &\qquad\times F(m_i) F(m_j) F(m_k) F(m_l).     
\end{split}
\q
We only consider the effect of merger history up to second-merger process.
Therefore the total merger rate density, $\cR(t, m_i, m_j)$, of PBH binaries at cosmic time $t$ with masses $m_i$ and $m_j$ is
\e\label{cR}
    \cR(t, m_i, m_j) =  \sum_{n=1,2} \cR_n(t, m_i, m_j),
\q 
and the total merger rate is
\e\label{Rt}
    R(t) = \int \cR(t, m_i, m_j) \rd m_i \rd m_j = \sum_{n=1,2} R_n(t),
\q 
where
\e 
 R_n(t) = \int \cR_n(t, m_i, m_j) \rd m_i \rd m_j.
\q 
All the above mentioned merger rate (density) is measured at source frame.
We should emphasize that although $R_2(t)$ should be smaller than $R_1(t)$ as expected, $\cR_2(t, m_i, m_j)$ is not necessarily be smaller than $\cR_1(t, m_i, m_j)$ \citep{Liu:2019rnx}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{method}Hierarchical Bayesian Inference}
To infer the population parameters, we adopt a hierarchical Bayesian approach by marginalizing over the uncertainty in the estimate of individual event parameters.
In this section, we describe the hierarchical Bayesian inference used in the parameter estimations.
The merger rate density \eqref{cR} is measured in source frame, and we need to convert it into the detector frame as
\e\label{Rpop}
\mR_{\mathrm{pop}}(\theta|\Ld) = \frac{1}{1+z} \frac{dV_\mathrm{c}}{dz} \mR(\theta|\Ld),
\q
where $z$ is the cosmological redshift, $\theta\equiv\{z, m_1, m_2\}$, $\Ld$ is a collection of $\fpbh$ and the parameters from mass function $P(m)$, and $dV_\mathrm{c}/dz$ is the differential comoving volume. The factor $1/(1 + z)$ converts time increments from the source frame to the detector frame. In this work, we take the cosmological parameters from Planck 2018 \citep{Planck:2018vyg}.

Given the data, $\textbf{d} = \{d_1, d_2, \cdots, d_{N_{\mathrm{obs}}}\}$, of $N_{\mathrm{obs}}$ BBH merger events, we model the total number of events as an inhomogeneous Poisson process, yielding the likelihood \citep{Loredo:2004nn,Thrane:2018qnx,Mandel:2018mve}
\begin{equation}\label{L1}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{\int \mathcal{L} (d_{i}| \theta)\, \cR_{\mathrm{pop}}(\theta|\Ld) d \theta}{\xi(\Ld)},
\end{equation}
where $N_{\exp}\equiv N_{\exp}(\Ld)$ is the expected number of detections over the timespan of observation, and $\mathcal{L} (d_{i}|\theta)$ is the individual event likelihood for the $i$th GW event that can be derived from the individual event's posterior by reweighing with the prior on $\theta$. Here, $\xi(\Ld)$ quantifies selection biases for a population with parameters $\Ld$ and is defined by
\e\label{xi}
\xi(\Ld) = \int P_{\mathrm{det}}(\theta)\, \cR_{\mathrm{pop}}(\theta|\Ld)\, \mathrm{d} \theta,
\q
where $P_{\text{det}}(\theta)$ is the detection probability that depends on the source parameters $\theta$.
In practice, we use the simulated injections \citep{ligo_scientific_collaboration_and_virgo_2021_5546676} to estimate $\xi(\Ld)$, and \Eq{xi} can be approximated by a Monte Carlo integral over found injections \citep{LIGOScientific:2021psn}
\begin{equation}
	\xi(\Ld) \approx \frac{1}{N_{\mathrm{inj}}} \sum_{j=1}^{N_{\text {found }}} \frac{\cR_{\mathrm{pop}}(\theta_{j} | \Ld)}{p_{\mathrm{draw}}(\theta_j)},
\end{equation}
where $N_{\text{inj}}$ is the total number of injections, $N_{\text{found}}$ is the number of successfully detected injections, and $p_{\mathrm{draw}}$ is the probability density function from which the injections are drawn.
Using the posterior samples from each individual event, the hyper-likelihood \eqref{L1} can be estimated as
\begin{equation}\label{L2}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{1}{\xi(\Ld)} \left\langle \frac{\cR_{\mathrm{pop}}(\theta|\Ld)}{d_L^2(z)} \right\rangle,
\end{equation}
where $\langle\cdots\rangle$ denotes the weighted average over posterior samples of $\theta$. The denominator $d_L^2(z)$ is the standard priors used in the LVK analysis of individual events where $d_L$ is the luminosity distance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[tbp!]
	\centering
	\begin{tabular}{lll}
		\hline\hline
		\textbf{Parameter\quad} & \textbf{Description} & \textbf{Prior} \\
		\hline
		$\fpbh$ & Abundance of PBH in dark matter & log-$\mU(-4, 0)$\\
		\hline
		\multicolumn{3}{c}{Lognormal PBH mass function} \\[1pt]
		$\Mc$ & Central mass in $\Msun$. & $\mU(5, 50)$\\
		$\sigma$ & Mass width. & $\mU(0.1, 2)$\\
		\hline
		\multicolumn{3}{c}{Power-law PBH mass function} \\[1pt]
		$\Mmin$ & Lower mass cut-off in $\Msun$. & $\mU(3, 10)$\\
		$\al$ & Power-law index. & $\mU(1.05, 4)$\\
		\hline
		\multicolumn{3}{c}{Broken Power-law PBH mass function} \\[1pt]
		$M_*$ & Peak mass in $\Msun$. & $\mU(3, 15)$\\
		$\al_1$ & First power-law index. & $\mU(0, 15)$\\
		$\al_2$ & Second power-law index. & $\mU(1, 3)$\\
		\hline
		\multicolumn{3}{c}{Critical collapse (CC) PBH mass function} \\[1pt]
		$\Mf$ & Horizon mass scale in $\Msun$. & $\mU(1, 50)$\\
		$\al$ & Universal exponent. & $\mU(0, 5)$\\
		\hline
	\end{tabular}	
	\caption{\label{table:priors}Parameters and their prior distributions used in the Bayesian parameter estimations. Here, $\mU$ and log-$\mU$ denote uniform and log-uniform distributions, respectively.}
\end{table}
Based on the hierarchical statistical framework, we do the parameter estimations for four different PBH mass functions that are commonly used in the literature. These mass functions are the log-normal, power-law, broken power-law, and critical collapse (CC) distributions, respectively. We summarize the parameters and their prior distributions in \Table{table:priors}. Below we show the results for each of the PBH mass function.

\subsection{Log-normal mass function}
We first consider a PBH mass function with the log-normal form \citep{Dolgov:1992pu} of
\e\label{log}
P(m) = \frac{1}{\sqrt{2\pi} \s m} \exp \(-\frac{\ln^2\(m/M_c\)}{2\s^2}\),
\q
where $M_c$ represents the central mass of $m P(m)$, and $\s$ characterizes the width of the mass spectrum.
A huge class of extended mass distributions can be approximated by the log-normal mass function if PBHs are formed from a smooth symmetric peak in the inflationary power spectrum when the slow-roll approximation holds \citep{Green:2016xgy,Carr:2017jsz,Kannike:2017bxn}.
The hyper-parameters are $\Ld = \{M_c, \s, \fpbh\}$ in this case. 
We can then derive the averaged PBH mass and averaged number density from \Eq{mpbh} and \Eq{Fm} as
\e
\mpbh = M_c \exp \(-\frac{\s^2}{2}\),
\q
\e 
F(m) = \frac{M_c}{\sqrt{2\pi} \s m^2} \exp \(-\frac{\s^2}{2}-\frac{\ln^2\(m/M_c\)}{2\s^2}\).
\q

@@

Using $69$ BBHs from GWTC-3 and performing the hierarchical Bayesian inference, we obtain $\Mf = 10.8^{+3.7}_{-3.6} \Msun$, $\al = 1.0^{+0.2}_{-0.2}$, and $\fpbh = 1.5^{+0.2}_{-0.2} \times 10^{-3}$. The posteriors for the hyper-parameters $\Ld = \{\Mf, \al, \fpbh\}$ are shown in \Fig{posterior-CC}. From \Eq{R0}, we also infer the local merger rate as $R_0 = 49^{+26}_{-16} \gpcyr$. The results of local merger rate and abundance of PBHs are consistent with the previous estimations \citep{Chen:2018czv,Chen:2018rzo,Chen:2019irf,Wu:2020drm,Chen:2021nxo,Chen:2022fda}, confirming that CDM cannot be dominated by the stellar-mass PBHs.

>>>>

%\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Primordial black holes (PBHs) can be not only cold dark matter candidates but also progenitors of binary black holes (BBHs) observed by LIGO-Virgo-KAGRA (LVK) Collaboration.
The PBH mass can be shifted to the heavy distribution if multi-merger processes occur.
In this work, we constrain the merger history of PBH binaries using the gravitational wave (GW) events from the third GW Transient Catalog (GWTC-3). Considering four commonly used PBH mass functions,   
namely the log-normal, power-law, broken power-law, and critical collapse forms, we find that the multi-merger processes make a subdominant contribution to the total merger rate. Therefore, the effect of merger history can be safely ignored when estimating the merger rate of PBH binaries. We also find that GWTC-3 is best fitted by the log-normal form among the four PBH mass functions and confirm that the stellar-mass PBHs cannot dominate cold dark matter.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The successful detection of gravitational waves (GWs) from compact binary coalescences \citep{LIGOScientific:2018mvr,LIGOScientific:2020ibl,LIGOScientific:2021djp} has led us into a new era of GW astronomy.
According to the recently released third GW Transient Catalog (GWTC-3) \citep{LIGOScientific:2021djp} by LIGO-Virgo-KAGRA (LVK) Collaboration, there are $90$ GW events detected during the first three observing runs. Most of these events are categorized as binary black hole (BBH) mergers, and the BBHs detected by LVK have a broad mass distribution. The heaviest event, GW190521 \citep{LIGOScientific:2020iuh}, has component masses $m_1 = 85^{+21}_{-14}\Msun$ and $m_2 = 66^{+17}_{-18}\Msun$. Both masses lie within upper black hole mass gap originated from pulsation pair-instability supernovae \citep{Marchant_2019}, and current modelling places the lower cutoff of the mass gap at $\sim 50\pm 4\Msun$ \citep{Marchant_2019,Farmer_2019,Farmer:2020xne,Marchant:2020haw}.
Even accounting for the uncertainties in modelling, it still implies at least $m_1$ is well within the mass gap and cannot have originated directly from a stellar progenitor \citep{Anagnostou:2020umw}.
Therefore, the heavy event GW190521 greatly challenges the stellar evolution scenario of astrophysical black holes.

Besides the astrophysical black holes, another possible explanation for the LVK BBHs is the primordial black holes (PBHs) \citep{Bird:2016dcv,Sasaki:2016jop,Chen:2018czv,Chen:2021nxo}. 
PBHs are black holes formed in the very early Universe through the gravitational collapse of the primordial density fluctuations \citep{Hawking:1971ei,Carr:1974nx}. The formation of PBHs would inevitably accompany the production of scalar-induced GWs \citep{Yuan:2019udt,Yuan:2019wwo,Yuan:2019fwv,Chen:2019xse,DeLuca:2019ufz,Bartolo:2018rku,Bartolo:2018evs}. PBHs can also be natural candidates of cold dark matter. Recent studies \citep{Chen:2021nxo,Chen:2022fda} show that the BBHs from GWTC-3 are consistent with the PBH scenario, and the abundance of PBH in CDM, $\fpbh$, should be in the order of $\mathcal{O}(10^{-3})$ to explain LVK BBHs. In particular, the merger rate for GW190521 derived from the PBH model is consistent with that inferred by LVK, indicating that GW190521 can be a PBH binary \citep{Chen:2021nxo}.

Accurately estimating the merger rate distribution of PBH binaries can be crucial to extract the PBH population parameters from GW data. \cite{Liu:2019rnx} study the multi-merger processes of PBH binaries and show that the merger history of PBH binaries may shift the mass distribution from light mass to heavy mass depending on the values of population parameters. \cite{Wu:2020drm} then infers the population parameters of PBH binaries by accounting for the merger history effect using $10$ BBHs from GWTC-1, finding that the effect of merger history can be safely ignored when estimating the merger rate of PBH binaries. 
In this work, we use the LVK recent released GWTC-3 data to constrain the effect of merger history on the merger rate of PBH binaries assuming all LVK BBHs are of primordial origin.
We extend the analyses of \cite{Wu:2020drm} in several aspects. Firstly, we use GWTC-3, which expands GWTC-1 with almost six times more BBH events. The GWTC-3 events expand the mass and redshift coverage and can alleviate the statistical bias by including significantly more BBHs. Secondly, \cite{Wu:2020drm} only consider the PBH mass functions with the log-normal and power-law forms. We do more comprehensive analyses by including the broken power-law and critical collapse PBH mass functions that were not considered in \cite{Wu:2020drm}. It is claimed by \cite{Deng:2021ezy} that a broken power-law can fit the GW data better than the log-normal form. Lastly, we consider the redshift distribution of the merger rate that was ignored in \cite{Wu:2020drm}.

We organize the rest paper as follows.
In \Sec{merger}, we briefly review the calculation of the merger rate of PBH binaries by accounting for the merger history effect.
In \Sec{method}, we describe the hierarchical Bayesian framework used to infer the PBH population parameters from GW data.
In \Sec{result}, we consider four commonly used PBH mass functions and present the results.
Finally, we give conclusions in \Sec{conclusion}.

@@

In this section, we will outline the calculation of merger rate density when considering the PBH merger history effect. We refer to \cite{Liu:2019rnx} for more details.

The BBHs observed by LVK suggest that BHs should have a broad mass distribution, so we consider an extended mass function for PBHs. Here, we demand the probability distribution function of PBH mass, $P(m)$, be normalized such that
\e\label{norm}
\int_{0}^{\infty} P(m)\, \rd m = 1.
\q 
Assuming the fraction of PBHs in CDM is $\fpbh$, we can estimate the abundance of PBHs in the mass interval $(m, m+\rd m)$ as \citep{Chen:2018rzo}
\e 
0.85 \fpbh\, P(m)\, \rd m.
\q 
The coefficient $0.85$ is roughly the fraction of CDM in the non-relativistic matter, including both CDM and baryons. Following \cite{Liu:2019rnx}, we may define an average PBH mass, $\mpbh$, as
\e\label{mpbh}
\frac{1}{\mpbh} = \int \frac{P(m)}{m} \rd m.
\q 
Then, we can obtain the average number density of PBHs with mass $m$ in the 
total number density of PBHs, $F(m)$, by \citep{Liu:2019rnx}
\e\label{Fm} 
F(m) = P(m) \frac{\mpbh}{m}.
\q

We can now estimate the merger rate densities of PBH binaries by considering the merger history effect.
We assume that PBHs are randomly distributed in the early Universe when they decouple from the cosmic background evolution \citep{Nakamura:1997sm,Sasaki:2016jop,Ali-Haimoud:2017rtz}. The two nearest PBHs would attract each other because of the gravitational interactions. These two PBHs would obtain the angular momentum from the torque of other PBHs and form a PBH binary. The binary would emit gravitational radiations and eventually merge.

We do not intend to give a detailed derivation but quote the results from \cite{Liu:2019rnx} here.
The merger rate density from first-merger process, $\cR_1(t, m_i, m_j)$, is given by \citep{Liu:2019rnx}
\e 
\cR_1(t, m_i, m_j) = \int \hat{\cR}_1 \rd m_l,
\q 
where
\e 
\begin{split}
	&\hat{\cR}_1(t, m_i, m_j, m_l)
	\equiv 1.32 \times 10^6 \times \(\frac{t}{t_0}\)^{-\frac{34}{37}}\(\frac{\fpbh}{\mpbh}\)^\frac{53}{37} \\ 
	&\times m_l^{-\frac{21}{37}} (m_i m_j)^\frac{3}{37} (m_i + m_j)^\frac{36}{37} F(m_i) F(m_j) F(m_l).     
\end{split}
\q
Similarly, the merger rate density from second-merger process, $\cR_2(t, m_i, m_j)$, is given by \citep{Liu:2019rnx}
\e 
\begin{split}
	\cR_2(t, m_i, m_j) &= \hf \int \hat{\cR}_2(t, m_i-m_e, m_e, m_j, m_l)\ \rd m_l \rd m_e\\
	&+ \hf \int \hat{\cR}_2(t, m_j-m_e, m_e, m_i, m_l)\ \rd m_l \rd m_e,
\end{split}
\q 
where
\e 
\begin{split}
	&\hat{\cR}_2(t, m_i, m_j,m_k, m_l)
	=1.59 \times 10^4 \times \(\frac{t}{t_0}\)^{-\frac{31}{37}} \(\frac{\fpbh}{\mpbh}\)^\frac{69}{37}\\ 
	&\qquad\times m_k^{\frac{6}{37}} m_l^{-\frac{42}{37}} (m_i+ m_j)^\frac{6}{37} (m_i + m_j + m_k)^\frac{72}{37}\\
	&\qquad\times F(m_i) F(m_j) F(m_k) F(m_l).     
\end{split}
\q
We only consider the effect of merger history up to the second-merger process.
Therefore the total merger rate density, $\cR(t, m_i, m_j)$, of PBH binaries at cosmic time $t$ with masses $m_i$ and $m_j$ is
\e\label{cR}
\cR(t, m_i, m_j) =  \sum_{n=1,2} \cR_n(t, m_i, m_j),
\q 
and the total merger rate is
\e\label{Rt}
R(t) = \int \cR(t, m_i, m_j) \rd m_i \rd m_j = \sum_{n=1,2} R_n(t),
\q 
where
\e 
R_n(t) = \int \cR_n(t, m_i, m_j) \rd m_i \rd m_j.
\q 
All the above-mentioned merger rate (density) is measured at the source frame.
We should emphasize that although $R_2(t)$ should be smaller than $R_1(t)$ as expected, $\cR_2(t, m_i, m_j)$ is not necessarily be smaller than $\cR_1(t, m_i, m_j)$ \citep{Liu:2019rnx}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{method}Hierarchical Bayesian Inference}
We adopt a hierarchical Bayesian approach to infer the population parameters by marginalizing the uncertainty in estimating individual event parameters.
This section describes the hierarchical Bayesian inference used in the parameter estimations.
The merger rate density \eqref{cR} is measured in the source frame, and we need to convert it into the detector frame as
\e\label{Rpop}
\mR_{\mathrm{pop}}(\theta|\Ld) = \frac{1}{1+z} \frac{dV_\mathrm{c}}{dz} \mR(\theta|\Ld),
\q
where $z$ is the cosmological redshift, $\theta\equiv\{z, m_1, m_2\}$, $\Ld$ is a collection of $\fpbh$ and the parameters from mass function $P(m)$, and $dV_\mathrm{c}/dz$ is the differential comoving volume. The factor $1/(1 + z)$ converts time increments from the source to the detector frame. We take the cosmological parameters from Planck 2018 \citep{Planck:2018vyg}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[tbp!]
	\centering
	\begin{tabular}{lll}
		\hline\hline
		\textbf{Parameter\quad} & \textbf{Description} & \textbf{Prior} \\
		\hline
		$\fpbh$ & Abundance of PBH in dark matter & log-$\mU(-4, 0)$\\
		\hline
		\multicolumn{3}{c}{Lognormal PBH mass function} \\[1pt]
		$\Mc$ & Central mass in $\Msun$. & $\mU(5, 50)$\\
		$\sigma$ & Mass width. & $\mU(0.1, 2)$\\
		\hline
		\multicolumn{3}{c}{Power-law PBH mass function} \\[1pt]
		$\Mmin$ & Lower mass cut-off in $\Msun$. & $\mU(3, 10)$\\
		$\al$ & Power-law index. & $\mU(1.05, 4)$\\
		\hline
		\multicolumn{3}{c}{Broken Power-law PBH mass function} \\[1pt]
		$M_*$ & Peak mass in $\Msun$. & $\mU(3, 15)$\\
		$\al_1$ & First power-law index. & $\mU(0, 15)$\\
		$\al_2$ & Second power-law index. & $\mU(1, 3)$\\
		\hline
		\multicolumn{3}{c}{Critical collapse (CC) PBH mass function} \\[1pt]
		$\Mf$ & Horizon mass scale in $\Msun$. & $\mU(1, 50)$\\
		$\al$ & Universal exponent. & $\mU(0, 5)$\\
		\hline
	\end{tabular}	
	\caption{\label{table:priors}Parameters and their prior distributions used in the Bayesian parameter estimations. Here, $\mU$ and log-$\mU$ denote uniform and log-uniform distributions, respectively.}
\end{table}

Given the data, $\textbf{d} = \{d_1, d_2, \cdots, d_{N_{\mathrm{obs}}}\}$, of $N_{\mathrm{obs}}$ BBH merger events, we model the total number of events as an inhomogeneous Poisson process, yielding the likelihood \citep{Loredo:2004nn,Thrane:2018qnx,Mandel:2018mve}
\begin{equation}\label{L1}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{\int \mathcal{L} (d_{i}| \theta)\, \cR_{\mathrm{pop}}(\theta|\Ld) d \theta}{\xi(\Ld)},
\end{equation}
where $N_{\exp}\equiv N_{\exp}(\Ld)$ is the expected number of detections over the timespan of observation. Here $\mathcal{L} (d_{i}|\theta)$ is the individual event likelihood for the $i$th GW event that can be derived from the individual event's posterior by reweighing with the prior on $\theta$. Here, $\xi(\Ld)$ quantifies selection biases for a population with parameters $\Ld$ and is defined by
\e\label{xi}
\xi(\Ld) = \int P_{\mathrm{det}}(\theta)\, \cR_{\mathrm{pop}}(\theta|\Ld)\, \mathrm{d} \theta,
\q
where $P_{\text{det}}(\theta)$ is the detection probability that depends on the source parameters $\theta$.
In practice, we use the simulated injections \citep{ligo_scientific_collaboration_and_virgo_2021_5546676} to estimate $\xi(\Ld)$, and \Eq{xi} can be approximated by a Monte Carlo integral over found injections \citep{LIGOScientific:2021psn}
\begin{equation}
	\xi(\Ld) \approx \frac{1}{N_{\mathrm{inj}}} \sum_{j=1}^{N_{\text {found }}} \frac{\cR_{\mathrm{pop}}(\theta_{j} | \Ld)}{p_{\mathrm{draw}}(\theta_j)},
\end{equation}
where $N_{\text{inj}}$ is the total number of injections, $N_{\text{found}}$ is the number of successfully detected injections, and $p_{\mathrm{draw}}$ is the probability density function from which the injections are drawn.
Using the posterior samples from each event, we estimate the hyper-likelihood \eqref{L1} as
\begin{equation}\label{L2}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{1}{\xi(\Ld)} \left\langle \frac{\cR_{\mathrm{pop}}(\theta|\Ld)}{d_L^2(z)} \right\rangle,
\end{equation}
where $\langle\cdots\rangle$ denotes the weighted average over posterior samples of $\theta$. The denominator $d_L^2(z)$ is the standard priors used in the LVK analysis of individual events where $d_L$ is the luminosity distance.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Critical collapse mass function}
